---
title: "Prereqs"
output:
  xaringan::moon_reader:
    css: ["default", "custom.css"]
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

class: dv

# Variables

* A way to store data in R

* Created using the _assignment operator_ `<-`

--

```{r}
x <- 5
```

* The variable `x` now stores the number 5

--

.pull-left[

* Print the contents of `x`

```{r}
x
```

]

--

.pull-right[

* Manipulate `x` as if it were `5`

```{r}
x * 10

x + 2
```

]

---

class: dv

# Data types

* There are different types of data in R

  * Numeric, Character (string), Boolean (True / False), Dates, etc
  
  * The "type" is also known as the "class" of data
  
--

.pull-left[

```{r}
# A numeric
my_number <- 10000
class(my_number)
```

]

--

.pull-right[

```{r}
# A character
# Notice the quotes
my_char <- "Hello!"
class(my_char)
my_char
```

]

---

class: dv

# Vectors

* Think of a vector like a _column in Excel_

  * Just a sequence of items / numbers
  
  * Each element in a vector is like a _cell in a column in Excel_
  
  * However, they can only hold 1 data type at a time (only numeric, only character)

--

.pull-left[

```{r}
# Vectors are created with
# the function: c()
policy_number <- c(5, 7, 87, 1000)

policy_number

policy_number / 2 
```

]

--

.pull-right[

```{r}
# You cannot mix 
# characters and numerics
# Numerics are forced as Character 
mixed_vector <- c("a string", 14)

# Notice the quotes around "14" 
mixed_vector 
```

]

---

class: dv

# Data frames / Tibbles

.pull-left[

* Why did I teach you all this so far?

* Building up to _data frames_

]

--

.pull-right[

* _Tables in Excel_

* Multiple columns, each column is a vector

]

--

```{r}
library(tibble)

df <- tibble(
  policy_name = c("Jason", "Danielle", "Andrew"),
  premium     = c(    120,      250.4,      300)
)

df
```

--

.center[Data frames are the main way to hold data in R]

---

class: dv

# Quick summary

* Variables - used to store data by name

* Vectors - multiple numbers / characters stored together

  * For us, mainly: numeric, character, dates
  
  * Only one data type per vector
  
* Data frames 

  * Holding multiple vectors together at once
  
  * Like _tables in Excel_
  
---

class: dv, center, middle

# Moving on... 

[Import](3-import.html)





