---
title: "Visualization"
output:
  xaringan::moon_reader:
    css: ["default", "custom.css"]
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, include = FALSE}
library(tidyverse)
library(nycflights13)
ggplot2::theme_set(new = ggplot2::theme_minimal())
```


class: dv

# Visualization

* The `ggplot2` package

  * Grammar of Graphics
  
  * Building a chart from layers
  
  * Incredibly flexible
  
---

class: dv

# 30 year treasury projections

* 3 scenarios

* 5 years of projections per scenario

```{r, eval=FALSE}
rates <- read_csv("data/example_rates3.csv")
rates
```

```{r, echo=FALSE, message = FALSE}
rates <- read_csv("data/example_rates3.csv") %>% 
  mutate(date = zoo::as.yearmon(date),
         scenario = as.character(scenario)) %>% 
  select(-asset) %>%
  rename(returns = return)
rmarkdown:::print.paged_df(rates)
```

---

class: dv

# ggplot()

* Every plot begins with a call to `ggplot()`

* This sets up a blank canvas for your charts

```{r, message=FALSE, fig.height=4, fig.width=9, fig.align='center', dpi=500, out.width="650px"}
ggplot()
```

--

.center[
## Thrilling, I know.
]

---

class: dv

# ggplot() data / aesthetics

* The data you are going to be charting is defined with `data =`

* The columns you map to the x-axis and y-axis are `aesthetics`

```{r, message=FALSE, fig.height=4, fig.width=9, fig.align='center', dpi=500, out.width="650px"}
ggplot(
  data    = rates, 
  mapping = aes(x = date, y = returns))
```

---

class: dv

# ggplot() +

* Add layers with `+`

* Layers like points or lines are called _geometries_. Add with `geom_point` or `geom_line`

```{r, message=FALSE, fig.height=4, fig.width=9, fig.align='center', dpi=500, out.width="650px"}
ggplot(
  data    = rates, 
  mapping = aes(x = date, y = returns)) +
{{ geom_point()}}
```

---

class: dv

# Grouping

* We have another dimension to look at, `scenario`

* Adding a `color` aesthetic will add a unique color for each scenario

```{r, message=FALSE, fig.height=4, fig.width=9, fig.align='center', dpi=500, out.width="650px"}
ggplot(
  data    = rates, 
  mapping = aes(x = date, y = returns, color = scenario)) +
  geom_point()
```


---

class: dv

# Geometries

* Line graph?

* Just switch out `geom_point` for `geom_line`

```{r, message=FALSE, fig.height=4, fig.width=9, fig.align='center', dpi=500, out.width="650px"}
ggplot(
  data    = rates, 
  mapping = aes(x = date, y = returns, color = scenario)) +
{{ geom_line()}}
```

---

class: dv

# Geometries

* Bar graph of median returns per scenario?

--

* Switch out `geom_point` for `geom_bar`. Scenario on x, Returns on y. Fill not color.


```{r, message=FALSE, fig.height=4, fig.width=9, fig.align='center', dpi=500, out.width="650px"}
ggplot(
  data    = rates, 
  mapping = aes(x = scenario, y = returns, fill = scenario)) +
{{ geom_bar(stat  = "summary", fun.y = "median") }}
```

---

class: dv

# Faceting

* Instead of 3 lines on 1 chart...

* ...can I put 1 line on 3 charts?

```{r, message=FALSE, fig.height=4, fig.width=9, fig.align='center', dpi=500, out.width="650px"}
ggplot(
  data    = rates, 
  mapping = aes(x = date, y = returns, color = scenario)) +
  geom_line()
```

---

class: dv

# Faceting

* Faceting allows you to split one chart into many

* `facet_wrap`

```{r, message=FALSE, fig.height=4, fig.width=9, fig.align='center', dpi=500, out.width="625px"}
ggplot(
  data    = rates, 
  mapping = aes(x = date, y = returns)) +
  geom_line() +
{{ facet_wrap(~scenario, nrow = 1) }}
```

---

class: dv

# Faceting

* Faceting allows you to split one chart into many

* `facet_wrap`

```{r, message=FALSE, fig.height=4, fig.width=9, fig.align='center', dpi=500, out.width="625px"}
ggplot(
  data    = rates, 
  mapping = aes(x = date, y = returns)) +
  geom_line() +
{{ facet_wrap(~scenario, nrow = 3) }}
```

---

class: dv

# Labels

* `labs()` to change the x-axis, y-axis, legend, title, ...

* Change the legend title by setting the label for the relevant aesthetic (`color`)

```{r, message=FALSE, fig.height=4, fig.width=9, fig.align='center', dpi=500, out.width="600px"}
ggplot(
  data    = rates, 
  mapping = aes(x = date, y = returns, color = scenario)) +
  geom_line() +
{{ labs(x = "Date", y = "Treasury Rate", color = "Scenario #", title = "Treasury rates over 3 scenarios", subtitle = "Projection period of 5 years", caption  = "Data from MG-ALFA - 08/01/2017")}}
```

---

class: dv

# Colors

* Manually set colors using `scale_color_manual`

* Integrated color palettes using `scale_color_brewer` from http://colorbrewer2.org


```{r, message=FALSE, fig.height=4, fig.width=9, fig.align='center', dpi=500, out.width="625px"}
ggplot(
  data    = rates, 
  mapping = aes(x = date, y = returns, color = scenario)) +
  geom_line() +
{{ scale_color_manual(values = c("#37424A", "#822433", "#B89002")) }}
```



---

class: dv, center, middle

# Next up..

[Manipulation](./5-manipulation.html)