---
title: "Database manipulation"
output:
  html_document:
    highlight: tango
    df_print: paged
---

```{r, message = FALSE, include = FALSE}
library(DBI)
library(dbplyr)
library(tidyverse)
```

### Connect to the database

```{r, warning=FALSE}
# Connect to database
cn <- dbConnect(drv = odbc::odbc(), 
                dsn = "fictrad")

# Last 6 tables
dbListTables(cn) %>% tail(6)
```

### Let's check out the 1q16 table

```{r}
# Connect to the 1q16 table
tbl_1q16 <- tbl(cn, "1q16")

tbl_1q16
```

### What about 3q16?

```{r}
# Connect to the 3q16 table
tbl_3q16 <- tbl(cn, "3q16")

tbl_3q16
```

---

## Formulate a question!

### _How did the Average Initial Statuatory Reserve change from 1q16 to 3q16?_

#### What if we break it down by `//ALFA_PLANCODE`, and by `ALFA_Sex`, and take the average `InitStatRes`.

---


### Adding a column of just "1q16"

```{r}
tbl_1q16 <- tbl_1q16 %>%
  
  # Add a quarter column
  mutate(quarter = "1q16") %>%
  
  # Only columns we are interested in for the question
  select(`//ALFA_PLANCODE`, ALFA_Sex, quarter, InitStatRes)
  
tbl_1q16
```

### Adding a column of just "3q16"

```{r}
tbl_3q16 <- tbl_3q16 %>%
  
  # Add a quarter column
  mutate(quarter = "3q16") %>%
  
  # Only columns we are interested in for the question
  select(`//ALFA_PLANCODE`, ALFA_Sex, quarter, InitStatRes)

tbl_3q16
```

### SQL is doing the hard work!

```{r}
tbl_1q16 %>% show_query()
```

### Actually pull them down from the database

```{r}
start_time <- proc.time()

# "Collect" the datasets from the database
tbl_1q16 <- collect(tbl_1q16)
tbl_3q16 <- collect(tbl_3q16)

proc.time() - start_time
```

### Stack the two datasets together. How large is this??

```{r}
# Bind the datasets together by rows
joined_tbl <- bind_rows(tbl_1q16, tbl_3q16)

# Dimensions
dim(joined_tbl)
```

### Average Statuatory Reserve from 1q16 -> 3q16

#### Broken down by plancode and sex 

```{r}
joined_tbl %>%
  
  # Grouping by plancode, sex, and quarter
  group_by(`//ALFA_PLANCODE`, ALFA_Sex, quarter) %>%
  
  # For each group, calculate the average stat reserve
  summarise(avg_stat_reserve = mean(InitStatRes)) %>%
  
  # Spread it out by quarter so you can compare the two
  spread(key = quarter, value = avg_stat_reserve)
```


