---
title: "Case Study - 1 - DeAlmeida"
author: "Davis Vaughan"
date: "9/6/2017"
output:
  html_document:
    df_print: paged
    highlight: tango
    theme: flatly
    code_folding: "show"
    code_download: "true"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Building a Life Table in R

```{r, message=FALSE, warning = FALSE}
library(tidyverse)
```

### Add ages

"Current industry mortality tables stop at age 121."

```{r}
# A vector 65, 66, ..., 121
ages_vector <- 65:121

# Create a "tibble". Basically a data frame.
# Columns are specified as `col_name = col_data`
life_table <- tibble(ages = ages_vector)

life_table
```

### Add the current year

We will also want to have a variable for the current year, `t`, starting at 0.

```{r}
life_table <- life_table %>%
  mutate(t = ages - 65)

life_table
```

### Add your p's and q's

"Assume 10 lives per 1000 mortality at age 65."

$$ q_{65} = 0.01 $$

$$ p_{65} = 0.99 $$

---

"The _force of mortality_ increases by 10% every year."

What does this mean? It means we could write the sequence of increasing 
force of mortality values as:

$$\begin{align} 
\mu_{x}   &= \mu \\ 
\mu_{x+1} &= \mu * 1.1 \\ 
\mu_{x+2} &= \mu * 1.1^2 
\end{align}$$

---

So $\mu$ is a constant...

And how do we relate force of mortality to probability of survival?

$$ p_x  = e ^ {-\int_{0}^{1} \mu_x dx} $$

For $p_{65}$, let $\mu_{65} = \mu$.

$$\begin{align}
p_{65} &= e ^ {-\int_{0}^{1} \mu  dx} \\ &=  e^{(-\mu*1) -(-\mu * 0)} \\ &= e ^ {-\mu} 
\end{align}$$

---

From $p_{65}$ to $p_{66}$, what happens? Force of mortality increases by 10%.


$$\begin{align} 
p_{66} &= e ^ {-1.1 \times\mu}  \\ &= (e ^ {-\mu} ) ^ {1.1}  \\ &= (p_{65}) ^ {1.1} 
\end{align}$$

---

From $p_{66}$ to $p_{67}$, what happens? Force of mortality increases by 10%.


$$\begin{align} 
p_{67} &= e ^ {-1.1^2 \times\mu} \\ &= (e ^ {-\mu} ) ^ {1.1^2} \\ &= (p_{65}) ^ {1.1^2}
\end{align}$$

---

The pattern emerges...

$$ p_{65} = (p_{65}) ^ {1.1^0} \\ p_{66} = (p_{65}) ^ {1.1^1} \\ p_{67} = (p_{65}) ^ {1.1^2} \\ ... \\ p_{65 + t} = (0.99) ^ {1.1^t}$$

---

$$ q_{65 + t} = 1 - p_{65 + t} $$

```{r}
life_table <- life_table %>%
  mutate(
    `p_65+t` = .99 ^ (1.1 ^ t),
    `q_65+t` = 1 - `p_65+t`
  )

life_table
```

### A visualization

It's worth checking out what we have done so far. Let's look at increasing 
mortality over time.

```{r, fig.align='center', fig.width=8.5}
life_table %>%
  
  # Renaming columns. These become the legend labels
  rename(p = `p_65+t`, q = `q_65+t`) %>%
  
  # Gather p and q columns together
  gather(key = p_or_q, value = prob, -ages, -t) %>%
  
  # The code that creates the plot
  ggplot(aes(x = ages, y = prob, color = p_or_q)) + 
  geom_line() +
  
  # Customization
  labs(x = "Current Age", y = "Probability", 
       title = "Probability of living (or dying) through 1 more year") +
  theme_minimal() +
  hrbrthemes::scale_color_ipsum()
```


### Adding $_{t}p_{65}$

We have $p_{65+t}$, how do we generate $_{t}p_{65}$ from this?

At first glance, this might seem difficult. In Excel we went back and forth 
between columns E and F a lot, and had the special case of $_{0}p_{65} = 1$ 
at the beginning.

It's totally possible to do this in R, but you would have to use a _for loop_ 
to implement that recursive calculation. Can we do something smarter?

---

Let's explore a few iterations of $_{t}p_{65}$.

$$\begin{align} 
_{0}p_{65} &= 1 \\ 
_{1}p_{65} &= p_{65 + 0} \\ 
_{2}p_{65} &= p_{65 + 0} \times p_{65 + 1} &= _{1}p_{65} \times p_{65 + 1} \\
_{3}p_{65} &= p_{65 + 0} \times p_{65 + 1} \times p_{65 + 2} &= _{2}p_{65} \times p_{65 + 2}
\end{align}$$

---

Notice that the middle equations there are a _cumulative product_ of $p_{65+t}$.

$$\begin{align} 
\text{original} &\rightarrow \text{cumprod}                                          &= _{t}p_{65} \\
1               &\rightarrow 1                                                       &= _{0}p_{65} \\ 
p_{65 + 0}      &\rightarrow 1 \times p_{65 + 0}                                     &= _{1}p_{65} \\ 
p_{65 + 1}      &\rightarrow 1 \times p_{65 + 0} \times p_{65 + 1}                   &= _{2}p_{65} \\ 
p_{65 + 2}      &\rightarrow 1 \times p_{65 + 0} \times p_{65 + 1} \times p_{65 + 2} &= _{3}p_{65} \\ 
\end{align}$$

---

Steps to accomplish this in R?

1) Lag (shift down) $p_{65+t}$
2) Fill in a 1 where we lost a value from the lag
3) Cumulative product of that column to generate $_{t}p_{65}$

---

```{r}
# An initial attempt at lagging p_{65+t}
life_table %>%
  mutate(
    p_lag = lag(`p_65+t`)
  )

# Lagging p_{65+t} but filling the NA with 1
life_table %>% 
  mutate(
    p_lag               = lag(`p_65+t`),
    p_lag_1             = lag(`p_65+t`, default = 1)
  )

# Now take the cumulative product of those lagged values
life_table %>% 
  mutate(
    p_lag               = lag(`p_65+t`),
    p_lag_1             = lag(`p_65+t`, default = 1),
    p_lag_cumprod       = lag(`p_65+t`, default = 1) %>% cumprod()
  )

# That was kind of ugly, let's round a bit
life_table %>% 
  mutate(
    p_lag               = lag(`p_65+t`),
    p_lag_1             = lag(`p_65+t`, default = 1),
    p_lag_cumprod       = lag(`p_65+t`, default = 1) %>% cumprod(),
    p_lag_cumprod_round = lag(`p_65+t`, default = 1) %>% cumprod() %>% round(digits = 6)
  )

# Now let's just do it all in one step and save it as `t_p_65`
life_table <- life_table %>% 
  mutate(
    `t_p_65` = lag(`p_65+t`, default = 1) %>% 
                  cumprod() %>% 
                  round(digits = 6)
  )

life_table
```

### Another visualization!

```{r, fig.align='center', fig.width=8.5}
life_table %>%
  
  # The code that creates the plot
  ggplot(aes(x = t, y = t_p_65)) + 
  geom_line() +
  
  # Customization
  labs(x = "Number of years, t, past age 65", y = "Probability of survival for t years", 
       title = "Probability of living for t years past age 65") +
  theme_minimal() +
  hrbrthemes::scale_color_ipsum()
```

### Finally, $l_{65 + t}$

Use the formula:

$$ _{t}p_{65} = \frac{l_{65+t}}{l_{65}} $$

---

Rearrange:

$$ l_{65+t}  = l_{65} \times _{t}p_{65} $$

---

* We know $l_{65} = 600$
* We calculated the column $_{t}p_{65}$ 

```{r}
# Original number of lives
l_65 = 600

# Calculate number of people alive at time t
life_table <- life_table %>%
  mutate(
    `l_65+t` = l_65 * t_p_65
  )

# Just round everything to 4 decimal places for pretty printing
life_table %>%
  modify(~round(.x, 4))
```

### One last visualization

```{r, fig.align='center', fig.width=8.5}
life_table %>%
  
  # The code that creates the plot
  ggplot(aes(x = t, y = `l_65+t`)) + 
  geom_line() +
  
  # Customization
  labs(x        = "Number of years, t, past age 65",
       y        = "Number of people alive at t years past 65", 
       title    = "Visualizing deaths at t years past 65", 
       subtitle = "Notice this has the same shape as the previous chart") +
  theme_minimal() +
  hrbrthemes::scale_color_ipsum()
```


### Without the explanation, how much code?

```{r}
# Parameters
initial_age  <- 65
ending_age   <- 121
l_x          <- 600      # 60mil / 100thousand
initial_q    <- 10/1000  # 10 lives out of 1000 at 65
f_o_m_factor <- 1.1      # from the 10% increase each year

# Age range
age_range <- initial_age:ending_age

# Initialize table with ages
life_table <- tibble(ages = age_range)

# Add all the other columns
life_table <- life_table %>%
  mutate(
    
    # Add the t years
    t = ages - initial_age,
    
    # p using the force of mortality factor
    `p_x+t` = (1 - initial_q) ^ (f_o_m_factor ^ t),
    
    # q is 1-p
    `q_x+t` = 1 - `p_x+t`,
    
    # t_p_x from p
    `t_p_x` = lag(`p_x+t`, default = 1) %>% cumprod(),
    
    # Number of lives left
    `l_x+t` = l_x * t_p_x
  )

life_table %>%
  modify(~round(.x, 4))
```
